---
title: "PS and IPTW"
author: "Sierra"
date: "2023-09-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libs
```{r, message=FALSE, warning=FALSE}
library(epiR)
library(labelled)
library(data.table)
library(sjlabelled)
library(ggeasy)
library(reshape2)
library(fastDummies)
library(smd)
library(lubridate)
library(survminer)
library(survival)
library(adjustedCurves)
library(jtools)
library(cobalt)
library(MatchIt)
library(WeightIt)
library(survey)
library(janitor)
library(tidyr)
library(summarytools)
library(haven)
library(here)
library(coxed)
library(tableone)
library(tidyverse)
library(ggplot2)
library(tidylog)
library(flextable)
library(gtsummary)
```
## Create date var
```{r}
today = Sys.Date()

```


## Load data
```{r}
#data_missing <- read_csv(here("Data", "brf-fakedata-missing_17AUG23.csv")) 

data = read_csv(here("Data","ind-ai-no-missing_28AUG23.csv")) 

key = read_csv(here("Data", "01-09-23_fakedata-label-key.csv"))

# make vectors of continuous and categorical vars

cat = c("id", "sex", "phu", "ster", "out_ai", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "exp_vax", "comp", 
"exp_flu", "bf", "seas", "gest", "itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")

con = c("age", "futime","bw")

bin = c("sex", "ster", "out_ai", "comorb", "matai", "del", "rur", "exp_vax", "comp", 
"exp_flu", "itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")

covs = c("phu", "ster", "comorb", "matai","itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom", "care", "par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf")

```

## Labelling data with labelled package following this tutorial https://www.pipinghotdata.com/posts/2022-09-13-the-case-for-variable-labels-in-r/
```{r}
# make frame of just old names and new names
meta = key %>%
  select(old.name, new.name) %>%
  filter(old.name != "preterm") # key cannot have variables that data does not have


# create a named vector with deframe
labels = meta %>%
  deframe()

# Assign labels using splice operator so that they are matched and order doesn't matter
# "Splicing causes each name of the list to become an argument name."
data.lab = data %>%
  set_variable_labels(!!!labels)

```


# Create data dictionary automatically
```{r}
dict = data.lab %>%
  generate_dictionary()
```

# gtsummary uses label names automatically
```{r}
t1 = data.lab %>%
  #mutate(across(c(bin), as.logical)) %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  tbl_summary(by=exp_vax,
              type = all_dichotomous() ~ "categorical") %>%
  add_overall() 

# add level names to data

show_header_names(t1)


t1 = t1 %>%
  modify_header(
  label = '**Characteristic**',
  stat_0 = '**Whole population** (unweighted) N = 229,753',
  stat_1 = '**No**, N = 122,679',
   stat_2 = '**Yes**, N = 107,074' ) %>%
  modify_spanning_header(
    c(stat_1, stat_2) ~ 
      "**2009 A/H1N1 influenza vaccination during pregnancy (unweighted)**")

t1

# %>%  as_flex_table()

"Distribution of baseline study population, overall and by 2009 A/H1N1 influenza vaccination status"
```


```{r}
data.lab %>%
  group_by( exp_vax) %>%
  summarize_all(
    N = n(), 
    pct = 
  )
```


```{r}
# Stick together a bunch of tabyls?

data.lab %>%
  tabyl(par, exp_vax) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns(position = "front")
```

# add to ggplot with ggeasy
Mutate will remove labels, so we re-add them lower in the pipeline
```{r}
data.lab %>%
  mutate(exp_vax = as.factor(exp_vax)) %>%
  set_variable_labels(!!!labels) %>%
  ggplot(aes( x=age, y=bw, color=exp_vax)) +
  geom_point() +
  easy_labs()
```
# make vars characters?
```{r}
d2.lab = data %>%
  mutate(across(all_of(cat), as.character)) %>%
  set_variable_labels(!!!labels)

# make dictionary and see that levels are now present
dict2 = d2.lab %>%
  generate_dictionary()

# try applying labels again
val_labels(d2.lab$sex) = c(sex.lab)

val_labels(d2.lab$sex,0)

```



## Another method using lists
https://lefkiospaikousis.netlify.app/posts/2020-12-23-a-workflow-with-labelled-data/
```{r}

# Create a labelled list where names are var names and values are var labels
lablist = var_label(data.lab)

# Pass the var name from the list to make informative ggplot titles
data %>%
  ggplot(aes(x=phu)) +
  geom_bar() +
  labs(x=lablist[["phu"]])
```
## for value labels
The format is (label = value): val_labels(v) <- c(yes = 1, no = 3, bug = 5) 
1. split key into separate df for each variable
2. make each df long so that old.val is a column and new.val is a column that correspond
3. write a function that makes vectors from the dfs
c("new.lvl" = "old.lvl", )

```{r}
# make key so that each value has its own row

key2 = key %>%
  separate_rows(old.lvl, new.lvl, sep=",") %>%
  mutate(old.lvl = as.numeric(old.lvl)) %>%
  filter(!is.na(old.lvl))

# Create new var that combines old and new lvls
tst = key2 %>%
  mutate(term = paste0(dQuote(new.lvl), sep="=", old.lvl))

# Combine into one row per variable and one vector each

tst = tst %>%
  select(-c(new.name, old.lvl, new.lvl)) %>%
  aggregate(term ~ old.name, toString)


# apply labels does not work. Maybe because you have to specify the levels at the same time?

val_labels(data.lab$sex) = c(sex.lab)

sex.lab = unlist(sex.lab)

val_labels(data.lab$sex) 
  
  
sex.lab = tst %>%
  filter(old.name == "sex") %>%
  select(term) 

val_labels(data.lab$sex)

levels = tst %>%
  deframe()

val_labels(data.lab$sex) = lvl$sex

lvl$sex

lvl = split(tst, f= tst$old.name)

lvl = lapply(lvl, function(x) x[-1])

sex.lab = 
```


```{r}
key3 = key %>%
  separate_rows(old.name)



xy.list <- as.list(as.data.frame(t(key)))

xy.list <- setNames(split(key, seq(nrow(key))), rownames(key))

xy.list2 = lapply(xy.list,as.list)


k3 = key %>%
  filter(!is.na(old.lvl)) %>%
  select(-c(new.name)) %>%
  purrr::transpose()


name.list = key %>%
  filter(!is.na(old.lvl)) %>%
  select(old.name)

# split df by var name
k4 = key %>%
  filter(!is.na(old.lvl)) %>% # get rid of continuous
  select(-c(new.name)) %>% # remove new name
 split(k4$old.name) 


names(k4)

# remove first element of each list (old.name)
k4 = lapply(k4, function(x) x[-1])

# we can select the variable lists now
k4$phu








test = key %>%
  filter(!is.na(old.lvl)) %>% # get rid of continuous
  select(-c(new.name, new.lvl)) %>% # remove new name
  pivot_wider(names_from = old.name,
              values_from = old.lvl)


row.names(key) = key$old.name







```

