---
title: "12AUG23 table1 with brfss data"
author: "Sierra"
date: "2023-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libs
```{r}
library(janitor)
library(tidyr)
library(summarytools)
library(haven)
library(here)
library(coxed)
library(tableone)
library(tidyverse)
library(ggplot2)
library(tidylog)
library(flextable)
library(gtsummary)
```

## Load data
```{r}
data_missing <- read_csv(here("Data", "brf-fakedata-missing_12AUG23.csv")) %>%
  mutate(phu = case_when(phu %in% 1:10 ~ 1,
                        phu %in% 11:20 ~ 2,
                        phu %in% 21:30 ~ 3,
                        phu %in% 31:40 ~ 4,
                        phu %in% 41:50 ~ 5,
                        phu %in% 51:70 ~ 6,
                        TRUE ~ NA_real_))

data = read_csv(here ("Data", "brf-fakedata-no-missing_12AUG23.csv")) %>%
  mutate(phu = case_when(phu %in% 1:10 ~ 1,
                        phu %in% 11:20 ~ 2,
                        phu %in% 21:30 ~ 3,
                        phu %in% 31:40 ~ 4,
                        phu %in% 41:50 ~ 5,
                        phu %in% 51:70 ~ 6,
                        TRUE ~ NA_real_))
```
## output missing data. 465473 rows x 23 vars
```{r}
#output data
write.csv(data_missing, file=here::here("Data","brf-fakedata-missing_17AUG23.csv"), row.names=FALSE)
```
## output non missing data. 229753 x 23 vars

```{r}
#output data
write.csv(data, file=here::here("Data","brf-fakedata-no-missing_17AUG23.csv"), row.names=FALSE)
```

## Explore data

The code below is from my prev exploration of the other fake data. Redo to match current data. 
```{r}
#peek at data
head(data)

#look at classes
lapply(data, class)

# summary of data. Look for categories with small numbers of people
stview(dfSummary(data))

```

## Plot things
```{r}

```

## Create table 1
Check dist of potential confounders stratified by exposure

```{r}

# Get col names from df
x <- colnames(data)

# print them as a vector
dput(x)

# make vectors of continuous and categorical vars

cat = c("id", "sex", "phu", "ster", "out_ai", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "exp_vax", "comp", 
"exp_flu", "bf", "seas", "gest")

con = c("age", "futime","bw")

# save ones I need to this vector (removed id, ageg, exp_flu, exp_vax, out_ai, failed,  )
vars = c("sex", "phu", "ster", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "comp", 
"bf", "age", "futime", "seas", "bw", "gest")

# create crude table
  
t1_crude = CreateTableOne(strata= "exp_vax", vars = vars, data=data, test=F)
print(t1_crude, smd=T)
# ster, comorb, care, par, smk, rur, quint, br all above 0.10

# check missing data set
t1_crude_miss = CreateTableOne(strata= "exp_vax", vars = vars, data=data_missing, test=F)
print(t1_crude_miss, smd=T)
  


```
# try gtsummary

"Each variable in the data frame has been assigned an attribute label (i.e. attr(trial$trt, "label") == "Chemotherapy Treatment") with the labelled package. These labels are displayed in the {gtsummary} output table by default. Using {gtsummary} on a data frame without labels will simply print variable names in place of variable labels; there is also an option to add labels later."

Makes a great table. Investigate why some SMDs are negative. 

```{r}
data %>%
  mutate(across(all_of(cat), as.factor)) %>%
  select (c("sex", "phu", "ster", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "comp", 
"bf", "age", "futime", "seas", "bw", "gest", "exp_vax")) %>%
  tbl_summary(by=exp_vax) %>%
   add_difference(everything() ~ "smd") %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="tbl1.docx")

```



