---
title: "PS and IPTW"
author: "Sierra"
date: "2023-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libs
```{r}
library(MatchIt)
library(WeightIt)
library(survey)
library(janitor)
library(tidyr)
library(summarytools)
library(haven)
library(here)
library(coxed)
library(tableone)
library(tidyverse)
library(ggplot2)
library(tidylog)
library(flextable)
library(gtsummary)
```

## Load data
```{r}
data_missing <- read_csv(here("Data", "brf-fakedata-missing_17AUG23.csv")) 

data = read_csv(here ("Data", "brf-fakedata-no-missing_17AUG23.csv")) 

# make vectors of continuous and categorical vars

cat = c("id", "sex", "phu", "ster", "out_ai", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "exp_vax", "comp", 
"exp_flu", "bf", "seas", "gest")

con = c("age", "futime","bw")
```


## Create propensity score model with log reg

From SAS code:

%let ps_vars=mom_ageD conc_season mult_birth obcomp2 obcomp3 obcomp7 pe abrup 
			 m_asthma m_diabetes m_heart m_hyper mathp12 mathp13 mathp14 mathp17 mathp18 mathp19 mathp20 mathp21 
			 mathp22 mathp23 mathp24 mathp25 mathp26 mathp27 steroids m_rural phu_region parity_cat m_incquint antenatal smoking_cat;
			 
```{r}
# model specification for maternal vaccination
# variables are the same ones from the SAS code plus maternal ai

# 1. estimate ps using glm
fit.ps = glm(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai), data=data, family=binomial) 
# get the predicted values  
data$ps = predict(fit.ps, type="response")

# apply the summary function to the two exposure groups
tapply(data$ps, data$exp_vax, summary)
```
# Assess the overlap of the distributions
by plotting them!

```{r}

# plots totally overlap? Maybe something went wrong.
data %>%
  mutate(exp_vax = as.factor(exp_vax)) %>%
  ggplot(aes(x=ps, group=exp_vax, color= exp_vax)) +
  geom_density() +
  ggtitle("Propensity Scores") +
  theme_light()

```

## Compute the weights

Stabilize = T

Mean of stabilized weights is ~ 1
```{r}
weights = weightit(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) +factor(matai),
                   data = data, 
                   estimand = "ATE",
                   stabilize = T,
                   method = "ps")

summary(weights$weights)
```
## Check balance in weighted pseudo-population
with SMD tables and love plots
Cobalt pulls out balanced statistics and flags which ones are above the level we specify (here, 0.10)

```{r}
require(cobalt)

bal.tab(weights, un = T, thresholds = c(m = .1))
# all of them are balanced now!

love.plot(weights)
ggsave("loveplot.png")
```

## survey package must be used to make weighted table one

```{r}
# make binary vars factors
fct_data = data %>%
  mutate(across(all_of(cat), as.factor))

# convert df to survey object
svy_data <-  survey::svydesign(
    id = ~id,
    weights = ~weights$weights,
    data = fct_data)


# only the vars included in the ps plus matai
svy_data %>%
  tbl_svysummary(
    by = exp_vax,
    include = c("phu", "ster", "comorb", "matai", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "matai")) %>%
  add_difference(everything() ~ "smd") %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="tbl1-ps.docx")



```

