---
title: "abstract work"
author: "Sierra"
date: "2023-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libs
```{r, message=FALSE, warning=FALSE}
library(epiR)
library(labelled)
library(table1)
library(data.table)
library(sjlabelled)
library(ggeasy)
library(reshape2)
library(fastDummies)
library(smd)
library(lubridate)
library(survminer)
library(survival)
library(adjustedCurves)
library(jtools)
library(cobalt)
library(MatchIt)
library(WeightIt)
library(survey)
library(janitor)
library(tidyr)
library(summarytools)
library(haven)
library(here)
library(coxed)
library(tableone)
library(tidyverse)
library(ggplot2)
library(tidylog)
library(flextable)
library(gtsummary)
```
## Create date var
```{r}
today = Sys.Date()

```


## Load data
```{r}
#data_missing <- read_csv(here("Data", "brf-fakedata-missing_17AUG23.csv")) 

data = read_csv(here("Data","ind-ai-no-missing_28AUG23.csv")) 

key = read_csv(here("Data", "01-09-23_fakedata-label-key.csv"))

# make vectors of continuous and categorical vars

cat = c("id", "sex", "phu", "ster", "out_ai", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "exp_vax", "comp", 
"exp_flu", "bf", "seas", "gest", "itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")

con = c("age", "futime","bw")

bin = c("sex", "ster", "out_ai", "comorb", "matai", "del", "rur", "exp_vax", "comp", 
"exp_flu", "itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")

covs = c("phu", "ster", "comorb", "matai", "care", "par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf")

ai = c("itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")
```



## Labelling data with labelled package following this tutorial https://www.pipinghotdata.com/posts/2022-09-13-the-case-for-variable-labels-in-r/
```{r}
# make frame of just old names and new names
meta = key %>%
  select(old.name, new.name) %>%
  filter(old.name != "preterm") # key cannot have variables that data does not have


# create a named vector with deframe
labels = meta %>%
  deframe()

# Assign labels using splice operator so that they are matched and order doesn't matter
# "Splicing causes each name of the list to become an argument name."
data.lab = data %>%
  set_variable_labels(!!!labels)

```


# Create data dictionary automatically
```{r}
dict = data.lab %>%
  generate_dictionary()
```

# explore data
### Note total number of moms and babies, data structure, percent missing, labels of missing values.
```{r}
str(data.lab)

stview(dfSummary(data.lab))
```

# raw table 1
```{r}
t1 = data.lab %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  tbl_summary(by=exp_vax,
              type = all_dichotomous() ~ "categorical") %>%
  add_overall() 

# add level names to data

show_header_names(t1)

# make it look better
t1 = t1 %>%
  modify_header(
  label = '**Characteristic**',
  stat_0 = '**Whole population** (unweighted) N = 229,753',
  stat_1 = '**No**, N = 122,679',
   stat_2 = '**Yes**, N = 107,074' ) %>%
  modify_spanning_header(
    c(stat_1, stat_2) ~ 
      "**2009 A/H1N1 influenza vaccination during pregnancy (unweighted)**")

t1

# %>%  as_flex_table()

"Distribution of baseline study population, overall and by 2009 A/H1N1 influenza vaccination status"


t1$meta_data$df_stats
```

# Remove those with missing exp_vax. Note how many are lost. 
### Compare summary to initial summary.
```{r}
data.vax = data.lab %>%
  filter(!is.na(exp_vax)) #remove missing exposures

stview(dfSummary(data.vax))
```

# Remove those with any missing PS covariates. Note how many are lost. 
### Compare summary to previous two summaries.
```{r}
data.vax.nm = data.vax %>%
  drop_na(covs) #remove missing PS covariates

stview(dfSummary(data.vax.nm))
```

# raw table 1
```{r}
t1a = data.vax.nm %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  tbl_summary(by=exp_vax,
              type = all_dichotomous() ~ "categorical") %>%
  add_overall() 

# add level names to data

show_header_names(t1a)

# make it look better
t1a = t1a %>%
  modify_header(
  label = '**Characteristic**',
  stat_0 = '**Whole population** (unweighted) N = 229,753',
  stat_1 = '**No**, N = 122,679',
   stat_2 = '**Yes**, N = 107,074' ) %>%
  modify_spanning_header(
    c(stat_1, stat_2) ~ 
      "**2009 A/H1N1 influenza vaccination during pregnancy (unweighted)**")

t1a
```

 
### Take the proportions from gtsummary and calculate categorical smds for them
```{r}
meta_tbl = t1a$meta_data %>% # grab the meta data to have
  unnest()

# make smds in the table info
tbl.bal = t1a$meta_data %>% # grab the meta data
  unnest() %>% # put the lists into one df
  select(variable, label, by, p, n, N, median, p25, p75) %>% #will I have any continuous vars?
  pivot_wider(names_from = by, values_from = c(n, N, median, p25, p75, p)) %>% #make cols of p_1 and p_0
  mutate(cat_smd=abs((p_1-p_0)/(sqrt(((p_1*(1-p_1)+p_0*(1-p_0)))/2)))) # calculate SMD



# just the proportions and SMDs
tbl = t1a$meta_data %>% # grab the meta data
  unnest() %>% # put the lists into one df
  filter(summary_type == "categorical" & !is.na(by)) %>% # get rid of the continuous vars and overall proportions
  select(variable, label, by, p) %>% 
  pivot_wider(names_from = by, values_from = p, names_prefix="p_") %>% #make cols of p_1 and p_0
  mutate(cat_smd=abs((p_1-p_0)/(sqrt(((p_1*(1-p_1)+p_0*(1-p_0)))/2)))) # calculate SMD

```



# SAS code
## SMDs for continuous
/* continuous variables */ 
 
*** step 1: compute the coefficients of weighted variance; 
data data1_weights2; 
	set &data.; 
	iptw_2=&iptw.**2; 
run; 
 
proc means data=data1_weights2 noprint; 
	class &exposure.; 
	var &iptw.; 
	output out=out_w sum(&iptw.)=sum_w sum(iptw_2)=sum_w2; 
run; 
 
/* for each continuous variable */ 
*** step 2: specify vardef=weight; 
proc means data=&data. vardef=weight noprint; 
	class &exposure.; 
	var &continuous.; 
	weight &iptw.; 
	output out=out_mean_var mean=mean var=var; 
run; 
 
*** step 3: compute the unbiased weighted variance; 
data w_var0 w_var1; 
	merge out_mean_var out_w(keep=&exposure. sum_w sum_w2); 
	by &exposure.; 
 
	w_var=((sum_w**2)/((sum_w**2)-sum_w2))*var; 
 
	if &exposure.=0 then output w_var0; 
	if &exposure.=1 then output w_var1; 
run; 
 
*rename variables for standardized difference calculations; 
proc sql; 
	create table temp as 
	select a.mean as m1, a.w_var as v1, b.mean as m2, b.w_var as v2 
	from w_var1 as a, 
	w_var0 as b; 
quit; 
 
*compute standardized mean difference using unbiased weighted variance; 
data num_var_std_diff&suffix.; 
	retain variable m1 v1 m2 v2 std_diff; 
	set temp; 
 
	std_diff=abs((m1-m2)/(sqrt((v1+v2)/2))); 
	variable="&continuous"; 
 
	label m1='Mean for exposed group' 
		  v1='Weighted variance for exposed group' 
		  m2='Mean for unexposed group' 
		  v2='Weighted variance for unexposed group'; 
run; 
 
/* categorical variables */ 
 
data cat_var_std_diff&suffix.; 
	length variable $30 levels $10 p_0 8 p_1 8 std_diff 8; 
	stop; 
run; 
 
```{r}

```



## Create propensity score model with log reg

From SAS code:

%let ps_vars=mom_ageD conc_season mult_birth obcomp2 obcomp3 obcomp7 pe abrup 
			 m_asthma m_diabetes m_heart m_hyper mathp12 mathp13 mathp14 mathp17 mathp18 mathp19 mathp20 mathp21 
			 mathp22 mathp23 mathp24 mathp25 mathp26 mathp27 steroids m_rural phu_region parity_cat m_incquint antenatal smoking_cat;
			 
```{r}
# model specification for maternal vaccination
# variables are the same ones from the SAS code plus maternal ai

# 1. estimate ps using glm
fit.ps = glm(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai),
             data=data.vax.nm, family=binomial) 

# get the predicted values  
data.vax.nm$ps = predict(fit.ps, type="response")

# apply the summary function to the two exposure groups
tapply(data.vax.nm$ps, data.vax.nm$exp_vax, summary)

# what am I looking to see here?
```

# Assess the overlap of the distributions
by plotting them!

```{r}

# plots totally overlap? Maybe something went wrong.
data.vax.nm %>%
  mutate(exp_vax = as.factor(exp_vax)) %>%
  ggplot(aes(x=ps, group=exp_vax, color= exp_vax)) +
  geom_density() +
  ggtitle("Propensity Scores") +
  theme_light()

```

## Compute the weights

Stabilize = T

Mean of stabilized weights is ~ 1
```{r}
weight.res = weightit(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai),
                   data = data.vax.nm, 
                   estimand = "ATE",
                   stabilize = T,
                   method = "ps")

summary(weight.res$weights)
```


## Check balance in weighted pseudo-population
with SMD tables and love plots
Cobalt pulls out balanced statistics and flags which ones are above the level we specify (here, 0.10)

```{r}
bal.tab(weight.res, un = T, thresholds = c(m = .1))
# all of them are balanced now!

love.plot(weight.res)
#ggsave("loveplot.png")
```


## Get SMDs so I can put them in a table
```{r}
d2 = data.vax.nm %>%
  mutate(across(all_of(cat), as.factor))

c2 = subset(d2, select = covs)
  

b = bal.tab(exp_vax ~ c2, data=data.vax.nm,
        weights = weight.res$weights,
        estimand="ATE",
        disp = c("means"), un = TRUE,
        stats = c("mean.diffs")) 

# Make the balance results a df
b1=b$Balance

# create a column with the rownames
b1$names <- rownames(b1)

# next, figure out how to stick it into t1 so that the right names align

```


## survey package must be used to make weighted table one (but actually I think I don't want a weighted table?)

```{r}
# make binary vars factors
fct_data = data.vax.nm %>%
  mutate(across(all_of(cat), as.factor))

# convert df to survey object
svy_data <-  survey::svydesign(
    id = ~id,
    weights = ~weight.res$weights,
    data = fct_data)


# only the vars included in the ps plus matai
tbl1_wght = svy_data %>%
  tbl_svysummary(
    by = exp_vax,
    include = c("phu", "ster", "comorb", "matai", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas")) %>%
  add_difference(everything() ~ "smd") %>%
  as_flex_table() 



tbl1_wght # only gives overall diffs

#tbl1_wght %>%
 # save_as_docx(path = here("Output", paste0(today,"_tbl1-weighted.docx")))
```


## Adjusted survival curves from adjustedCurves package
```{r}

# outcome must be numeric or logical but exposure must be factor
curve = data.vax.nm %>%
  mutate(exp_vax = as.factor(exp_vax))

# create log model

glm_mod = glm(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai) +factor(itp)+factor(jia)+factor(t1d)+factor(ibd)+factor(celiac)+factor(vasc)+factor(sjo)+factor(glom),
                   data = curve, 
                   family = "binomial" (link="logit"))

# calc adjusted survival curves and plot them

adjsurv = adjustedsurv(data=curve, 
                       variable = "exp_vax",
                       ev_time = "futime",
                       event = "out_ai",
                       method = "iptw_cox",
                       treatment_model=glm_mod,
                       conf_int = T)

plot(adjsurv)
```



## Cox PH
```{r}
# crude Cox PH model
cox_model_crude <- coxph(Surv(futime, out_ai) ~ exp_vax, 
                         data = data.vax.nm)
# get HR estimates
cox_fit_crude  <- cox_model_crude %>% 
  broom::tidy(., conf.int = T, exponentiate = T)

#summary 
summary(cox_model_crude)

# IPTW Cox PH model
cox_model_iptw <- coxph(Surv(futime, out_ai) ~ exp_vax, 
                        data = data.vax.nm, 
                        weights = weight.res$weights)
# get HR estimates
cox_fit_iptw  <- cox_model_iptw %>% 
  broom::tidy(., conf.int = T, exponentiate = T)

#summary

summary(cox_model_iptw)

# Put HRs into a table. Top is crude, bottom is IPTW-adjusted
cox_fit_crude %>% 
  bind_rows(cox_fit_iptw)

```

# Try to make results table
### needs outcome, number of events, IR and 95% CI per 1000 PY by exposure status, crude estimate, and adjusted estimate

itp, jia, t1d, ibd, celiac, vasc, sjo, glom

```{r}
# get overall futime by exposure

data.vax.nm %>%
  select(futime, exp_vax) %>%
  group_by(exp_vax) %>%
  summarize(totalfu = sum(futime), totalfu.yr = (sum(futime))/365.25)

# trying just one outcome
 data.vax.nm %>%
  select (exp_vax, futime, t1d) %>%
  group_by(exp_vax) %>%
  summarize(event = sum(t1d), total.fu.yrs = (sum(futime))/365.25) %>%
  mutate(IR = (event/total.fu.yrs)*1000)
 
 
# count number of each disease and sum of futime by exposure status
outcomes = data.vax.nm %>%
  select(exp_vax, futime, itp, jia, t1d, ibd, celiac, vasc, sjo, glom) %>%
  group_by(exp_vax) %>%
  summarize(across(everything(), ~ sum(.x))) %>% #only add the 1s of the diseases
  mutate(futime.yrs = futime/365.25)
 
 
# Get table in the correct format
incidence.tbl = outcomes %>%
  pivot_longer(!c(exp_vax, futime, futime.yrs), names_to = "ai_disease", values_to = "events") %>%
  mutate(IR = (events/futime.yrs)*1000) %>%
  pivot_wider(names_from = exp_vax, 
              values_from=c(futime, futime.yrs, events, IR),
              names_vary = "slowest")
```



