---
title: "abstract work"
author: "Sierra"
date: "2023-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libs
```{r, message=FALSE, warning=FALSE}
library(epiR)
library(labelled)
library(table1)
library(data.table)
library(sjlabelled)
library(ggeasy)
library(reshape2)
library(fastDummies)
library(smd)
library(lubridate)
library(survminer)
library(survival)
library(adjustedCurves)
library(jtools)
library(cobalt)
library(MatchIt)
library(WeightIt)
library(survey)
library(janitor)
library(tidyr)
library(summarytools)
library(haven)
library(here)
library(coxed)
library(tableone)
library(tidyverse)
library(ggplot2)
library(tidylog)
library(flextable)
library(gtsummary)
```
## Create date var
```{r}
today = Sys.Date()

```


## Load data
```{r}
#data_missing <- read_csv(here("Data", "brf-fakedata-missing_17AUG23.csv")) 

data = read_csv(here("Data","ind-ai-no-missing_28AUG23.csv")) 

key = read_csv(here("Data", "01-09-23_fakedata-label-key.csv"))

# make vectors of continuous and categorical vars

cat = c("id", "sex", "phu", "ster", "out_ai", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "exp_vax", "comp", 
"exp_flu", "bf", "seas", "gest", "itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")

con = c("age", "futime","bw")

bin = c("sex", "ster", "out_ai", "comorb", "matai", "del", "rur", "exp_vax", "comp", 
"exp_flu", "itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")

covs = c("phu", "ster", "comorb", "matai", "care", "par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf")



dat = head(data, 300)

stview(dfSummary(d2))


d2 = data[sample(nrow(data), 300), ]

write.csv(d2, "D:/DocumentsHDD/uOttawa/Thesis/R coding/fake-thesis/FakeData/Data/data-stub.csv", row.names=FALSE)

str(data)

str(d2)
```



## Labelling data with labelled package following this tutorial https://www.pipinghotdata.com/posts/2022-09-13-the-case-for-variable-labels-in-r/
```{r}
# make frame of just old names and new names
meta = key %>%
  select(old.name, new.name) %>%
  filter(old.name != "preterm") # key cannot have variables that data does not have


# create a named vector with deframe
labels = meta %>%
  deframe()

# Assign labels using splice operator so that they are matched and order doesn't matter
# "Splicing causes each name of the list to become an argument name."
data.lab = data %>%
  set_variable_labels(!!!labels)

```


# Create data dictionary automatically
```{r}
dict = data.lab %>%
  generate_dictionary()
```

# explore data
```{r}
stview(dfSummary(data.lab))
```


```{r}
table(data.lab$sex, data.lab$exp_vax)

with(data.lab, ctable(sex, exp_vax))

str(data.lab)

data.lab %>%
  tabyl(sex, exp_vax)

%>%
  purrr::imap(.f = ~tabyl(dat = humans, !!sym(.y), gender) %>%
                     adorn_percentages("row") %>%
                     adorn_pct_formatting(digits = 2) %>%
                     adorn_ns())


data %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  purrr::imap(.f = ~janitor::tabyl(dat = data, !!sym(.y), exp_vax) %>% # !! unquotes, sym is symbol
                adorn_percentages("col") %>%
                adorn_pct_formatting(digits = 2) %>%
                adorn_ns())
```

```{r}
humans <- starwars %>%
    filter(species == "Human")

str(humans)

t2 <- humans %>%
    tabyl(gender, eye_color)

t2 %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns()

humans %>%
  select(where(is.character)) %>%
  select(-name, -gender) %>%
  purrr::imap(.f = ~tabyl(dat = humans, !!sym(.y), gender) %>%
                     adorn_percentages("row") %>%
                     adorn_pct_formatting(digits = 2) %>%
                     adorn_ns())
```

```{r}
t3 <- head(data)
  
  
t3 = t3 %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  mutate(exp_vax = factor(exp_vax)) %>%
    tabyl(par, exp_vax)

t3 %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns()

data %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  mutate(exp_vax = factor(exp_vax)) %>%
  purrr::imap(.f = ~tabyl(dat = data, !!sym(.y), exp_vax) %>%
                     adorn_percentages("row") %>%
                     adorn_pct_formatting(digits = 2) %>%
                     adorn_ns())


str(data)

class(data) <- c("tbl_df", "tbl", "data.frame")
```




# raw table 1
```{r}
t1 = data.lab %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  tbl_summary(by=exp_vax,
              type = all_dichotomous() ~ "categorical") %>%
  add_overall() 

# add level names to data

show_header_names(t1)

# make it look better
t1 = t1 %>%
  modify_header(
  label = '**Characteristic**',
  stat_0 = '**Whole population** (unweighted) N = 229,753',
  stat_1 = '**No**, N = 122,679',
   stat_2 = '**Yes**, N = 107,074' ) %>%
  modify_spanning_header(
    c(stat_1, stat_2) ~ 
      "**2009 A/H1N1 influenza vaccination during pregnancy (unweighted)**")

t1

# %>%  as_flex_table()

"Distribution of baseline study population, overall and by 2009 A/H1N1 influenza vaccination status"


t1$meta_data$df_stats
```

# Hilcode 
### Take the proportions from gtsummary and calculate categorical smds for them
```{r}
meta_tbl = t1$meta_data %>% # grab the meta data
  unnest()

tst = t1$meta_data$df_stats[[2]]

m2 = meta_tbl %>%
  left_join(tbl, by=c("variable", "label")) # join smds to meta data

m3 = m2 %>%
  select(var_label, by, label, median, p25, p75, N_obs, n, N, p, col_name) %>%
   pivot_wider(names_from = col_name, values_from = p, names_prefix="p_") %>% #make cols of p_1 and p_0
  mutate(cat_smd2=abs((p_stat_2-p_stat_1)/(sqrt(((p_stat_2*(1-p_stat_2)+p_stat_1*(1-p_stat_1)))/2)))) # calculate SMD

# totally works
tbl = t1$meta_data %>% # grab the meta data
  unnest() %>% # put the lists into one df
  filter(summary_type == "categorical" & !is.na(by)) %>% # get rid of the continuous vars and overall proportions
  select(variable, label, by, p) %>% 
  pivot_wider(names_from = by, values_from = p, names_prefix="p_") %>% #make cols of p_1 and p_0
  mutate(cat_smd=abs((p_1-p_0)/(sqrt(((p_1*(1-p_1)+p_0*(1-p_0)))/2)))) # calculate SMD

t1

# 
m4 = t1$meta_data %>% # grab the meta data
  unnest() %>% # put the lists into one df
  select(variable, label, by, p, n, N, median, p25, p75) %>% 
  pivot_wider(names_from = by, values_from = c(n, N, median, p25, p75, p)) %>% #make cols of p_1 and p_0
  mutate(cat_smd=abs((p_1-p_0)/(sqrt(((p_1*(1-p_1)+p_0*(1-p_0)))/2)))) # calculate SMD

#summarize option, stopped early

# either widen the data frame so that each level of the cat variables are columns, or
# pivot longer to have a column called var name/characteristic and a column called level ( multiple rows of each var) 
# group by both exp_vax and characteristic variable

data.lab %>%
  select(age, par, smk, comorb, matai, comp, multip, quint, rur, gest, sex, exp_vax) %>%
  group_by(exp_vax) %>%
  summarize(across(everything(), list(n=length, mean=mean)))

t1

# make summary table a tibble
t1.tbl = as_tibble(t1)

# join key to tbl to get new names with the data
name.tbl = tbl %>%
  left_join(key, by = c("variable" = "old.name"))

# join tbl with summary based on new.name
name.tbl %>%
  select (new.name, label, cat_smd) %>%
  full_join(key, by = c("new.name" = "variable", ))

```

```{r}
tst = t1$meta_data$df_stats[[2]]

tst %>%
  filter(!is.na(by)) %>%
  group_by(variable_levels) %>%
  

```



# SMDs for continuous
/* continuous variables */ 
 
*** step 1: compute the coefficients of weighted variance; 
data data1_weights2; 
	set &data.; 
	iptw_2=&iptw.**2; 
run; 
 
proc means data=data1_weights2 noprint; 
	class &exposure.; 
	var &iptw.; 
	output out=out_w sum(&iptw.)=sum_w sum(iptw_2)=sum_w2; 
run; 
 
/* for each continuous variable */ 
*** step 2: specify vardef=weight; 
proc means data=&data. vardef=weight noprint; 
	class &exposure.; 
	var &continuous.; 
	weight &iptw.; 
	output out=out_mean_var mean=mean var=var; 
run; 
 
*** step 3: compute the unbiased weighted variance; 
data w_var0 w_var1; 
	merge out_mean_var out_w(keep=&exposure. sum_w sum_w2); 
	by &exposure.; 
 
	w_var=((sum_w**2)/((sum_w**2)-sum_w2))*var; 
 
	if &exposure.=0 then output w_var0; 
	if &exposure.=1 then output w_var1; 
run; 
 
*rename variables for standardized difference calculations; 
proc sql; 
	create table temp as 
	select a.mean as m1, a.w_var as v1, b.mean as m2, b.w_var as v2 
	from w_var1 as a, 
	w_var0 as b; 
quit; 
 
*compute standardized mean difference using unbiased weighted variance; 
data num_var_std_diff&suffix.; 
	retain variable m1 v1 m2 v2 std_diff; 
	set temp; 
 
	std_diff=abs((m1-m2)/(sqrt((v1+v2)/2))); 
	variable="&continuous"; 
 
	label m1='Mean for exposed group' 
		  v1='Weighted variance for exposed group' 
		  m2='Mean for unexposed group' 
		  v2='Weighted variance for unexposed group'; 
run; 
 
/* categorical variables */ 
 
data cat_var_std_diff&suffix.; 
	length variable $30 levels $10 p_0 8 p_1 8 std_diff 8; 
	stop; 
run; 
 
```{r}

```








# trying to make a reprex for stack overflow question
```{r}
t1$meta_data$stat_display

x = head(data.lab, 30)

age = x$age

care = x$care

sex = x$sex

exp = x$exp_vax

id = c(1:30)

age

paste(exp, collapse = ', ')

tabyl(care)


df = data.frame(age, sex, care, exp)


tbl = df %>%
  tbl_summary(by=exp,
              type = all_dichotomous() ~ "categorical") %>%
  add_overall() 

tbl
```






```{r}
tab = table1(~ age+ factor(par)+ factor(smk)+ factor(comorb)+factor(matai)+ factor(comp)+ factor(multip)+ factor(quint)+ factor(rur)+ factor(gest)+ factor(sex)|  exp_vax , data=data.lab )
```


## Create propensity score model with log reg

From SAS code:

%let ps_vars=mom_ageD conc_season mult_birth obcomp2 obcomp3 obcomp7 pe abrup 
			 m_asthma m_diabetes m_heart m_hyper mathp12 mathp13 mathp14 mathp17 mathp18 mathp19 mathp20 mathp21 
			 mathp22 mathp23 mathp24 mathp25 mathp26 mathp27 steroids m_rural phu_region parity_cat m_incquint antenatal smoking_cat;
			 
```{r}
# model specification for maternal vaccination
# variables are the same ones from the SAS code plus maternal ai

# 1. estimate ps using glm
fit.ps = glm(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai),
             data=data.lab, family=binomial) 

# get the predicted values  
data.lab$ps = predict(fit.ps, type="response")

# apply the summary function to the two exposure groups
tapply(data.lab$ps, data.lab$exp_vax, summary)

# what am I looking to see here?
```
# Assess the overlap of the distributions
by plotting them!

```{r}

# plots totally overlap? Maybe something went wrong.
data.lab %>%
  mutate(exp_vax = as.factor(exp_vax)) %>%
  ggplot(aes(x=ps, group=exp_vax, color= exp_vax)) +
  geom_density() +
  ggtitle("Propensity Scores") +
  theme_light()

```

## Compute the weights

Stabilize = T

Mean of stabilized weights is ~ 1
```{r}
weight.res = weightit(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai),
                   data = data.lab, 
                   estimand = "ATE",
                   stabilize = T,
                   method = "ps")

summary(weight.res$weights)
```


## Check balance in weighted pseudo-population
with SMD tables and love plots
Cobalt pulls out balanced statistics and flags which ones are above the level we specify (here, 0.10)

```{r}
bal.tab(weight.res, un = T, thresholds = c(m = .1))
# all of them are balanced now!

love.plot(weight.res)
#ggsave("loveplot.png")
```

## Get SMDs so I can put them in a table
```{r}
d2 = data.lab %>%
  mutate(across(all_of(cat), as.factor))

c2 = subset(d2, select = covs)
  

b = bal.tab(exp_vax ~ c2, data=data.lab,
        weights = weight.res$weights,
        estimand="ATE",
        disp = c("means"), un = TRUE,
        stats = c("mean.diffs")) 

# Make the balance results a df
b1=b$Balance

# create a column with the rownames
b1$names <- rownames(b1)

# next, figure out how to stick it into t1 so that the right names align

```


## survey package must be used to make weighted table one (but actually I think I don't want a weighted table)

```{r}
# make binary vars factors
fct_data = data.lab %>%
  mutate(across(all_of(cat), as.factor))

# convert df to survey object
svy_data <-  survey::svydesign(
    id = ~id,
    weights = ~weight.res$weights,
    data = fct_data)


# only the vars included in the ps plus matai
tbl1_wght = svy_data %>%
  tbl_svysummary(
    by = exp_vax,
    include = c("phu", "ster", "comorb", "matai", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas")) %>%
  add_difference(everything() ~ "smd") %>%
  as_flex_table() 



tbl1_wght # only gives overall diffs

#tbl1_wght %>%
 # save_as_docx(path = here("Output", paste0(today,"_tbl1-weighted.docx")))
```


## Adjusted survival curves from adjustedCurves package
```{r}

# outcome must be numeric or logical but exposure must be factor
d2 = data %>%
  mutate(exp_vax = as.factor(exp_vax))

# create log model

glm_mod = glm(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai) +factor(itp)+factor(jia)+factor(t1d)+factor(ibd)+factor(celiac)+factor(vasc)+factor(sjo)+factor(glom),
                   data = d2, 
                   family = "binomial" (link="logit"))

# calc adjusted survival curves and plot them

adjsurv = adjustedsurv(data=d2, 
                       variable = "exp_vax",
                       ev_time = "futime",
                       event = "out_ai",
                       method = "iptw_cox",
                       treatment_model=glm_mod,
                       conf_int = T)

plot(adjsurv)
```



## Cox PH
```{r}
# crude Cox PH model
cox_model_crude <- coxph(Surv(futime, out_ai) ~ exp_vax, 
                         data = data)
# get HR estimates
cox_fit_crude  <- cox_model_crude %>% 
  broom::tidy(., conf.int = T, exponentiate = T)

#summary 
summary(cox_model_crude)

# IPTW Cox PH model
cox_model_iptw <- coxph(Surv(futime, out_ai) ~ exp_vax, 
                        data = data, 
                        weights = weights$weights)
# get HR estimates
cox_fit_iptw  <- cox_model_iptw %>% 
  broom::tidy(., conf.int = T, exponentiate = T)

#summary

summary(cox_model_iptw)

# Put HRs into a table
cox_fit_crude %>% bind_rows(cox_fit_iptw)

```
## Another surival plot

This one is not right
```{r}
ggsurvplot(survfit(cox_model_iptw), 
           data=data,
           ggtheme = theme_minimal())
```


# trying to combine cobalt and gtsummary into one table
```{r}
# categorical: std_diff=abs((p_1-p_0)/(sqrt(((p_1*(1-p_1)+p_0*(1-p_0))))/2))

tbl_regression_ex <-
  lm(age ~ grade + marker, trial) %>%
  tbl_regression() %>%
  bold_p(t = 0.5)

tbl_regression_ex

tbl_summary_ex <-
  trial %>%
  select(trt, age, grade, response) %>%
  tbl_summary(by = trt)

tbl_summary_ex

tbl_regression_ex %>%
  purrr::pluck("table_body")

tbl_summary_ex %>%
  purrr::pluck("table_body")

# can I make a table with my results, add a column for variable, add a col for label, and join on them by matching both?

body = t1 %>%
  purrr::pluck("table_body") %>%
  rename(stat_overall = stat_0,
         stat_no = stat_1,
         stat_yes = stat_2)

t1

b
  
```
# trying to combine gtsummary and cobalt SMDs into one table
```{r}

## Get SMDs so I can put them in a table

d2 = data.lab %>%
  mutate(across(all_of(cat), as.factor))

c2 = subset(d2, select = covs)
  

b = bal.tab(exp_vax ~ c2, data=data.lab,
        weights = weight.res$weights,
        estimand="ATE",
        disp = c("means"), un = TRUE,
        stats = c("mean.diffs")) 

# Make the balance results a df
b1 = b$Balance 

# create a column with the rownames
b1$names <- rownames(b1)

# create var and label columns for joining
b2 = b1 %>%
  select (names, Diff.Un, Diff.Adj) %>%
  mutate(var = sub("_.*", "", names),
         label = sub(".*_", "", names))


# join with summary body

left_join(body,
          b2,
          by = variable)


# matches on label and var, so binary variables are out
body %>%
  rename(var = variable) %>%
  left_join(b2)




library(unglue)
body %>%
  mutate(pct_no = unglue::unglue_vec(stat_no, 
  "{}( {x} %{}"))


body %>%
  mutate(pct_no = gsub(".*[(] ([^.]+) [%].*", "\\1", stat_no))

library(qdap)

body %>%
  mutate(n_no =  bracketX(stat_no))
```


```{r}
gtsummary::trial %>%
  select(response, age, marker) %>%
  tbl_uvregression(
    y = response,
    method = glm,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    hide_n = TRUE)



 tbl_custom_summary(
    include = c("Age", "Class"),
    by = "Sex",
    stat_fns = ~ proportion_summary("Survived", "Yes", weights = "Freq"),
    statistic = ~"{prop}% ({n}/{N}) [{conf.low}-{conf.high}]",
    digits = ~ list(
      function(x) {
        style_percent(x, digits = 1)
      },
      0, 0, style_percent, style_percent
    ),
    overall_row = TRUE,
    overall_row_last = TRUE
  ) %>%
  bold_labels() %>%
  modify_footnote(
    update = all_stat_cols() ~ "Proportion (%) of survivors (n/N) [95% CI]"
  )
```


```{r}
# get percent alone 

body %>%
   mutate(pct_no = stringr::str_extract(string = stat_no,
                                               pattern = "(?<=\\().*(?=\\))"),
          pct_yes = stringr::str_extract(string = stat_yes,
                                               pattern = "(?<=\\().*(?=\\))"))



v=-10

y = abs(v)


p_1 = runif(100, 0, 1)
p_0 = runif(100, 0 ,1)

std_diff=abs(
  (p_1-p_0)/
    (sqrt(((p_1*(1-p_1)
            +p_0*(1-p_0))))/2))


tbl
```

```{r}
# create data
library(tidyverse)
library(gtsummary)

age = c(34, 26, 42, 39, 17, 32, 28, 35, 42, 42, 44, 21, 26, 19, 34, 38, 20, 23, 33, 39, 28, 30, 19, 30, 37, 22, 32, 26, 18, 26)

care = c(4, 2, 4, 4, 5, 4, 4, 4, 4, 4, 4, 2, 4, 4, 4, 4, 4, 4, 5, 4, 4, 4, 4, 4, 2, 2, 4, 4, 4, 4)

sex = c(1, 1, 1, 1, 0, 1, 1, 1, 0, 1, 1, 1, 0, 1, 1, 0, 1, 0, 0, 0, 1, 0, 0, 1, 1, 0, 0, 1, 0, 1)

exp = c(1, 0, 1, 0, 0, 1, 0, 1, 0, 0, 0, 0, 1, 1, 0, 1, 0, 1, 0, 1, 1, 0, 1, 1, 0, 1, 0, 0, 1, 0)

df = data.frame( age, sex, care, exp)

# create table summary
tbl = df %>%
  tbl_summary(by=exp,
              type = all_dichotomous() ~ "categorical") %>%
  add_overall() 

tbl

# categorical SMD calculation, where p_1 is the proportion in exp group 1 and p_0 is the proportion in exp group 0 (the "by" variable). 

cat_smd=abs((p_1-p_0)/(sqrt(((p_1*(1-p_1)+p_0*(1-p_0))))/2))

library(reprex)
reprex()

p_1 = 0.5
p_0=0.31
```

