---
title: "Untitled"
author: "Sierra"
date: "2023-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libs
```{r, message=FALSE, warning=FALSE}
library(epiR)
library(sqldf)
library(msm)
library(lubridate)
library(survminer)
library(survival)
library(adjustedCurves)
library(jtools)
library(cobalt)
library(MatchIt)
library(WeightIt)
library(survey)
library(janitor)
library(tidyr)
library(summarytools)
library(haven)
library(here)
library(coxed)
library(tableone)
library(tidyverse)
library(ggplot2)
library(tidylog)
library(flextable)
library(gtsummary)
```
## Create date var
```{r}
today = Sys.Date()

# create function to draw from a restricted normal dist
trunc_norm <- function(mean, sd, lower, upper) {
  rtnorm(n = 1, mean = mean, sd = sd, lower = lower, upper = upper)
}

```


## Load data
```{r}
data_missing <- read_csv(here("Data", "brf-fakedata-missing_17AUG23.csv")) 

data = read_csv(here ("Data", "brf-fakedata-no-missing_17AUG23.csv")) 

# make vectors of continuous and categorical vars

cat = c("id", "sex", "phu", "ster", "out_ai", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "exp_vax", "comp", 
"exp_flu", "bf", "seas", "gest")

con = c("age", "futime","bw")

```

## 02 exposure script
```{r}
# spine of baby IKNs and dobs
spine = a_flu.finalcohort %>%
  select (m_ikn, b_ikn, m_key, mom_iknd, lmp)

# check that the maternal ikns match
spine %>%
  filter(m_ikn != mom_iknd) %>%
  view()

# create list of unique maternal ikns
mikn = spine %>%
  select(unique(m_ikn))
```

# create primary exposure

proc sql; 
	create table temp as 
	select a.flagH1N1collect, b.b_ikn, b.m_ikn, b.m_key, b.vaccination_influenza 
	from a_flu.indicateH1N1collect as a  
		left join a_flu.finalcohortLW as b 
	on a.b_ikn=b.b_ikn; 
quit; 
```{r}
flu.anyH1N1_vacc = a_flu.indicateH1N1collect %>%
  left_join(a_flu.finalcohort, by=b_ikn)%>%
  select(flagH1N1collect, b_ikn, m_ikn, m_key, vaccination_influenza) %>%
  mutate(invacdata = if_else(flagH1N1collect ==1, 1, 0),
         anyh1n1 = case_when(is.na(vaccination_influenza) ~ NA_real_,
                             vaccination_influenza %in% c(1, 2, 3, 4) ~ 1,
                             TRUE ~ 0),
         onlyh1n1 = case_when(is.na(vaccination_influenza) ~ NA_real_,
                             vaccination_influenza == 1 ~ 1,
                             TRUE ~ 0),
         anytiv = case_when(is.na(vaccination_influenza) ~ NA_real_,
                             vaccination_influenza %in% c(2, 3) ~ 1,
                             TRUE ~ 0),
         finalanyh1n1 = case_when(invacdata == 1 & is.na(vaccination_influenza) ~ NA_real_,
                                  invacdata == 1 & vaccination_influenza %in% c(1, 3, 4) ~ 1, #may be wrong
                             TRUE ~ 0)) %>%
  select(m_ikn, m_key, b_ikn, anyh1n1, onlyh1n1, anytiv, invacdata, finalanyh1n1)

# NOTE: H1N1 vaccination exposure variable going forward will be finalanyH1N1 (must exclude all women with birth outside of exposure collection period and with missing primary exposure information -- invac=0 and finalanyH1N1=.)

anyH1N1_vacc_subset = flu.anyH1N1_vacc %>%
  select(b_ikn, m_ikn, m_key, finalanyH1N1) %>%
  filter(!is.na(finalanyH1N1) & invacdata ==1)

# exclude infants who died on birthday (0 follow up)

subset2 = flu.hc_use %>%
  filter(b.tot_fu != 0) %>%
  inner_join(flu.anyH1N1_vacc_subset, by=b_ikn)

flu.anyH1N1_vacc_subset = subset2

```
# Secondary exposure of interest during pandemic influenza season using DAD, NACRS, & OHIP;

I skipped the code where the analyst got info from DAD because I don't think I will have to do that

```{r}
m_cohort = flu.infection_cohort %>%
  select(m_ikn, m_key, lmp, b_ikn, b_bdate) %>%
  rename(ikn = m_ikn) # new = old or vice versa?
```


