---
title: "PS and IPTW"
author: "Sierra"
date: "2023-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libs
```{r, message=FALSE, warning=FALSE}
library(epiR)
library(labelled)
library(data.table)
library(sjlabelled)
library(reshape2)
library(fastDummies)
library(smd)
library(lubridate)
library(survminer)
library(survival)
library(adjustedCurves)
library(jtools)
library(cobalt)
library(MatchIt)
library(WeightIt)
library(survey)
library(janitor)
library(tidyr)
library(summarytools)
library(haven)
library(here)
library(coxed)
library(tableone)
library(tidyverse)
library(ggplot2)
library(tidylog)
library(flextable)
library(gtsummary)
```
## Create date var
```{r}
today = Sys.Date()

```


## Load data
```{r}
#data_missing <- read_csv(here("Data", "brf-fakedata-missing_17AUG23.csv")) 

data = read_csv(here("Data","ind-ai-no-missing_28AUG23.csv")) 

key = read_csv(here("Data", "01-09-23_fakedata-label-key.csv"))

# make vectors of continuous and categorical vars

cat = c("id", "sex", "phu", "ster", "out_ai", "comorb", "matai", "care", 
"par", "smk", "del", "multip", "rur", "quint", "exp_vax", "comp", 
"exp_flu", "bf", "seas", "gest", "itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom")

con = c("age", "futime","bw")

covs = c("phu", "ster", "comorb", "matai","itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom", "care", "par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf")


```
## Df Summary
```{r}
stview(dfSummary(data))
```

## Create smd formula to calculate it for each level, not the overall variable
```{r}
lvlsmd <- function(data, variable, by, tbl, ...) {
  smd(data[[variable]] ~ as.factor(data[[by]]))
}


t1 = data %>%
  mutate(across(all_of(cat), as.factor)) %>%
  select (c("phu", "ster", "comorb", "matai","itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf", "exp_vax")) %>%
  tbl_summary(by=exp_vax) %>%
  add_stat(fns = everything() ~ SMD_value)
```
```{r}
SMD_value <- function(x, ...) {
  # Construct vectors of data y, and groups (strata) g
  y <- unlist(x)
  g <- factor(rep(1:length(x), times=sapply(x, length)))
  if (is.numeric(y)) {
    # For numeric variables
    try({a<-data.frame(y)
    a$g<-g
    smd<-(as.data.frame(stddiff.numeric(data=a,gcol = "g", vcol = "y")))$stddiff
    },silent=TRUE)
  } else {
  # For categorical variables
    try({
      a<-data.frame(y)
      a$g<-g
      smd<-(abs((bal.tab(a, treat = "g",data=a,binary="std",continuous =         
 "std",s.d.denom = "pooled",stats=c("mean.diffs"))$Balance)$Diff.Un))
    },silent=TRUE)
  }
  c("",format(smd,digits=2)) #Formatting number of digits
}

```


```{r}

d2 = data %>%
  mutate(across(all_of(cat), as.factor)) %>%
  select (c("phu", "ster", "comorb", "matai","itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf", "exp_vax")) %>%
  fastDummies::dummy_cols()


# tbl summary will give me them for each level if I make them all dummy vars
d2 %>%
  select (-c("phu", "ster", "comorb", "matai","itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom", "care", "par", "smk", "multip", "rur", "quint", "comp", "seas", "gest", "sex", "del", "bf", "exp_vax")) %>%
  tbl_summary(by=exp_vax) %>%
   add_difference(everything() ~ "smd") %>%
  as_flex_table()
```



## Crude table 1
```{r}
data %>%
  mutate(across(all_of(cat), as.factor)) %>%
  select (c("phu", "ster", "comorb", "matai","itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf", "exp_vax")) %>%
  tbl_summary(by=exp_vax) %>%
  add_overall() %>%
  modify_spanning_header(c(estimate)  ~ "**2009 A/H1N1 influenza vaccination in pregnancy (unweighted)**") 

#%>%   as_flex_table()



t1 = data %>%
  mutate(across(all_of(cat), as.factor)) %>%
  select (c("phu", "ster", "comorb", "matai","itp", "jia", "t1d", "ibd", "celiac", "vasc", "sjo", "glom", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas", "bw", "gest", "sex", "del", "bf", "exp_vax")) %>%
  tbl_summary(by=exp_vax) %>%
  add_overall() %>%
  modify_spanning_header(c(stat_1, stat_2)  ~ "**2009 A/H1N1 influenza vaccination in pregnancy (unweighted)**") 

#%>%   as_flex_table()

show_header_names(t1)

t1
```

## Hmisc to relabel things
```{r}
### Explore this as way to label variables and levels
# https://cran.r-project.org/web/packages/Hmisc/Hmisc.pdf
FEV2 <- upData(FEV,
rename=c(smoking='smoke'),
levels=list(sex =list(female=0,male=1),
smoke=list('non-current smoker'=0,
'current smoker'=1)),
units=list(age='years', fev='L', height='inches'),
labels=list(fev='Forced Expiratory Volume'))
# An alternative to levels=list(\dots) is for example
# upData(FEV, sex=factor(sex,0:1,c('female','male'))).
#
# Note that we saved the changed data frame into a
# new data frame FEV2. If we were confident of the
# correctness of our changes we could have stored
# the new data frame on top of the old one, under
# the original name FEV.

```



## Create propensity score model with log reg

From SAS code:

%let ps_vars=mom_ageD conc_season mult_birth obcomp2 obcomp3 obcomp7 pe abrup 
			 m_asthma m_diabetes m_heart m_hyper mathp12 mathp13 mathp14 mathp17 mathp18 mathp19 mathp20 mathp21 
			 mathp22 mathp23 mathp24 mathp25 mathp26 mathp27 steroids m_rural phu_region parity_cat m_incquint antenatal smoking_cat;
			 
```{r}
# model specification for maternal vaccination
# variables are the same ones from the SAS code plus maternal ai

# 1. estimate ps using glm
fit.ps = glm(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai) +factor(itp)+factor(jia)+factor(t1d)+factor(ibd)+factor(celiac)+factor(vasc)+factor(sjo)+factor(glom),
             data=data, family=binomial) 
# get the predicted values  
data$ps = predict(fit.ps, type="response")

# apply the summary function to the two exposure groups
tapply(data$ps, data$exp_vax, summary)


```
# Assess the overlap of the distributions
by plotting them!

```{r}

# plots totally overlap? Maybe something went wrong.
data %>%
  mutate(exp_vax = as.factor(exp_vax)) %>%
  ggplot(aes(x=ps, group=exp_vax, color= exp_vax)) +
  geom_density() +
  ggtitle("Propensity Scores") +
  theme_light()

```

## Compute the weights

Stabilize = T

Mean of stabilized weights is ~ 1
```{r}
weights = weightit(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai) +factor(itp)+factor(jia)+factor(t1d)+factor(ibd)+factor(celiac)+factor(vasc)+factor(sjo)+factor(glom),
                   data = data, 
                   estimand = "ATE",
                   stabilize = T,
                   method = "ps")

summary(weights$weights)
```
## Check balance in weighted pseudo-population
with SMD tables and love plots
Cobalt pulls out balanced statistics and flags which ones are above the level we specify (here, 0.10)

```{r}
bal.tab(weights, un = T, thresholds = c(m = .1))
# all of them are balanced now!

d2 = data %>%
  mutate(across(all_of(cat), as.factor))

c2 = subset(d2, select = covs)
  

b = bal.tab(exp_vax ~ c2, data=data,
        weights = weights$weights,
        estimand="ATE",
        disp = c("means"), un = TRUE,
        stats = c("mean.diffs")) 

# Make the balance results a df
b1=b$Balance

# create a column with the rownames
b1$names <- rownames(b1)


love.plot(weights)
#ggsave("loveplot.png")

```

```{r}


balpop = bal.tab(weights)

balpop$Balance

```



## survey package must be used to make weighted table one

```{r}
# make binary vars factors
fct_data = data %>%
  mutate(across(all_of(cat), as.factor))

# convert df to survey object
svy_data <-  survey::svydesign(
    id = ~id,
    weights = ~weights$weights,
    data = fct_data)


# only the vars included in the ps plus matai
tbl1_wght = svy_data %>%
  tbl_svysummary(
    by = exp_vax,
    include = c("phu", "ster", "comorb", "matai", "care", 
"par", "smk", "multip", "rur", "quint", "comp", "age", "seas")) %>%
  add_difference(everything() ~ "smd") %>%
  as_flex_table() 

tbl1_wght

tbl1_wght %>%
  save_as_docx(path = here("Output", paste0(today,"_tbl1-weighted.docx")))
```


## Adjusted survival curves from adjustedCurves package
```{r}

# outcome must be numeric or logical but exposure must be factor
d2 = data %>%
  mutate(exp_vax = as.factor(exp_vax))

# create log model

glm_mod = glm(exp_vax ~ age + factor(seas) + factor(multip) + factor(comp) + factor(comorb) + factor(ster) + factor(rur) + factor(phu) + factor(par) + factor(quint) + factor(care) + factor(smk) + factor(matai) +factor(itp)+factor(jia)+factor(t1d)+factor(ibd)+factor(celiac)+factor(vasc)+factor(sjo)+factor(glom),
                   data = d2, 
                   family = "binomial" (link="logit"))

# calc adjusted survival curves and plot them

adjsurv = adjustedsurv(data=d2, 
                       variable = "exp_vax",
                       ev_time = "futime",
                       event = "out_ai",
                       method = "iptw_cox",
                       treatment_model=glm_mod,
                       conf_int = T)

plot(adjsurv)
```
## Cox PH
```{r}
# crude Cox PH model
cox_model_crude <- coxph(Surv(futime, out_ai) ~ exp_vax, 
                         data = data)
# get HR estimates
cox_fit_crude  <- cox_model_crude %>% 
  broom::tidy(., conf.int = T, exponentiate = T)

#summary 
summary(cox_model_crude)

# IPTW Cox PH model
cox_model_iptw <- coxph(Surv(futime, out_ai) ~ exp_vax, 
                        data = data, 
                        weights = weights$weights)
# get HR estimates
cox_fit_iptw  <- cox_model_iptw %>% 
  broom::tidy(., conf.int = T, exponentiate = T)

#summary

summary(cox_model_iptw)

# Put HRs into a table
cox_fit_crude %>% bind_rows(cox_fit_iptw)

```
## Another surival plot

This one is not right
```{r}
ggsurvplot(survfit(cox_model_iptw), 
           data=data,
           ggtheme = theme_minimal())
```

